\begin{acknowledgements}

It is an immense privilege to be born in a society with means and resolve
to generously support education and science. Specifically, I have benefitted
from superb education provided by II. gimnazija Maribor, Imperial College
London, and University of Oxford, supported by the Slovenian and British
public. Similarly, the scientific content of this thesis critically depends
on the work of my ATLAS collaborators and colleagues at CERN operating the
LHC, all supported by the human populace around the world.

I am indebted in particular to individuals who devote their time and energy
to pass their knowledge to others, and do so with inspiration and enthusiasm.
I thank Barbara Pe\'canac, Zdenka Keuc, and Gorazd
\v{Z}iberna for their classes that have inspired countless students to pursue
further education in science. I am grateful to Adam Dobbs and Jordan Nash for
directing my interest to particle physics, and to Gavin Davies, Patrick Dunne,
and Bj\"{o}rn Penning for their patience and perseverance in directing my
first research efforts in high energy physics.

This thesis would not have been completed without the mentorship and guidance
of Ian Shipsey and Giacomo Artoni. I thank Ian for his enduring faith in my
abilities and the encouragement to tackle challenges that seemed insurmountable.
He taught me that it is better to set the goals high and never reach them,
than to set them too low and achieve them.
I thank Giacomo for the many hours of his time, spent on teaching me physics and
programming, and for the outstanding leadership of the analysis team. He taught
me that everyone has something to contribute and the importance of discovering 
and nurturing it.

I thank Max Bellomo, Stefano Rosati, and Federico Sforza for welcoming me to
the MCP group, and Arthur Lesage for his invaluable help with existing
software. I also thank Giacomo Artoni, Jan Kretzschmar, Haifeng Li, and Stefano
Rosati for their leadership of the $\hmumu$ analysis team, and the $\hmumu$
analysis contributors for their simultaneously creative and scientifically
rigorous work, for their relentless drive to push the sensitivity of the
available dataset to its absolute limits, and for the countless entertaining
meetings.

I am grateful to Daniela Bortoletto and Ian Shipsey for fostering an energetic
and inclusive particle physics group, and to Sue Geddes, Jennifer Matthews,
and Kim Proudfoot for their help with departmental administration. I am
thankful to Peter Braam, Sue Geddes, and Andrey Ustyuzhanin for the
collaboration on the MLHEP summer school and the ML+Physics seminar series.
I am also thankful to Pio Monti for his wise words on statistics and machine
learning, and to Vip Davda, Pete Gronbech, and Kashif Mohammad for their help
with computing.

Discussions with Luca Ambroz, Lydia Beresford, Mikkel Bj{\o}rn, Maria Giovanna
Foti, Tereza Kroupov\'a, Jesse Liu, Luigi Marchese, Nurfikri Norjoharuddeen,
Santiago Paredes, Mariyan Petrov, Wouter Van de Pontseele, Hannah Pullen,
Elisabeth Schopf, Beojan Stanislaus, Cecilia Tosciri, Iza Veli\v{s}\v{c}ek,
Yingjie Wei, Ricardo W\"olker, Philipp Windischhofer, and Siyuan Yan made me a
better physicist and a more broad-minded person, and I thank them for teaching
me about their subjects and cultures. I am particularly grateful to Nurfikri
for his guidance in the early days of the DPhil, to Philipp for intelectually
stimulating weekend discussions, and to Jesse and Luca for their enduring
companionship.

For keeping me sane and cheerful during the DPhil journey I thank Jan
Alati\v{c}, Vigor Arva, Miha Damjan, Lina Dobnikar, Gaja Golija, Xavier
Gonzalez, Miha Hom\v{s}ak, Jaka Kukavica, Vid Lepej, Bo\v{s}tjan Muminovi\'{c},
Nejc Pulko, Nina Purg, Uro\v{s} Vadlja, Martin Vogrin, Rok Vogrin, and Tadej
Volj\v{c}. They have been the finest of friends.

Finally, I thank Alenka, Bojan, Ines, Jure, and stara mama Nu\v{s}a for
trying to teach me it is usually more important to be nice than to be right,
for providing me with the opportunity to receive outstanding education, and
for their unwavering encouragement and support which enabled the completion
of this thesis.

\end{acknowledgements}

%\thispagestyle{empty}
